---
title: "Fraud-Analysis"
author: "Alvaro Pena"
date: "2023-01-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(xgboost)
library(caret)
library(Matrix)
library(data.table)
library(GGally)
library(DiagrammeR)
```

# Read data

```{r}
# .csv file can be obtained from the link in the readme section
data = read.csv2("data.csv", header = T, sep = ",")
```

Transforming type of data
```{r}
data$step = as.numeric(data$step)
data$type = as.factor(data$type)
data$amount = as.double(data$amount)
data$oldbalanceOrg = as.double(data$oldbalanceOrg)
data$newbalanceDest = as.double(data$newbalanceDest)
data$oldbalanceDest = as.double(data$oldbalanceOrg)
data$newbalanceOrig = as.double(data$newbalanceOrig)

data$isFraud = as.logical(data$isFraud)
data$isFlaggedFraud = as.logical(data$isFlaggedFraud)
```

```{r}
summary(data)
```
```{r}
count = data %>% select(type, isFraud) %>% filter(isFraud == T) 

nrow(count)
# We know now that frauds are comitted by taking out cash and tranfers
count %>% group_by(type) %>% count()
```

Fraudulent transactions occur when taking cash out and transferring money. It will be interesting to show if the type of transaction has any effect when classifying the transactions.

```{r}
emptied_accounts = data %>% filter(newbalanceOrig == 0)

summary(emptied_accounts)
```

```{r}
data %>% select(amount, isFraud) %>%
  ggplot(aes(x = log(amount))) +
  geom_density() +
  facet_wrap("isFraud") +
  theme_bw()
```
Fraudulent transaction handle higher sums of money than real one.

Stratified sampling using the caret package

Training data
```{r}
partition = createDataPartition(data$isFraud, p = 0.7, list = F)

training = data[partition, ] 

training.data = training %>% select(-c(isFraud, step, nameOrig, nameDest, isFlaggedFraud))
training.labels = training %>% select(isFraud) %>%
  mutate(isFraud = as.numeric(isFraud)) %>% pull()
```

Testing data
```{r}
test = data[-partition, ]

test.data = test %>% select(-c(isFraud, step, nameOrig, nameDest, isFlaggedFraud)) 
test.labels = test %>% select(isFraud) %>% mutate(isFraud = as.numeric(isFraud))
```

Making sure the proportion of fraudulent transactions hold for both sets
```{r}
sum(training$isFraud)/ nrow(training) * 100
sum(test$isFraud)/ nrow(test) * 100
```

We are creating sparse matrix for the xgboost model
```{r}
training_matrix = sparse.model.matrix(~.-1, training.data)

test_matrix = sparse.model.matrix(~.-1, test.data)
```

Training the model
```{r}
bst = xgboost(data = training_matrix, label = training.labels, max.depth = 2,
               eta = 1, nthread = 4, nrounds = 20,
              objective = "binary:logistic", verbose =  2)
```


```{r}
importance <- xgb.importance(feature_names = training_matrix@Dimnames[[2]], model = bst)
head(importance)

importanceRaw <- xgb.importance(feature_names = training_matrix@Dimnames[[2]], model = bst, data = training_matrix, label = training.labels)

# Cleaning for better display
importanceClean <- importanceRaw[,`:=`(Cover=NULL, Frequency=NULL)]

head(importanceClean)

xgb.plot.importance(importance_matrix = importanceRaw)

```
From our analysis the balance on the destination account and the balance of the origin account after the transaction are the most important
features. Surprisingly the type of transaction is not a relevant feature.

```{r}
str(data)
```

```{r}
pred = predict(bst, test_matrix)
prediction = as.numeric(pred > 0.5)

test$isFraudPred = prediction

# This is the same as doing
# error = mean(test$isFraudPred != as.numeric(test$isFraud)) * 100

error = sum(test$isFraudPred != as.numeric(test$isFraud))/ nrow(test) * 100
```
Error is very small 0.04% of transactions are not classified as fraudulent. Out of 191 thousand transactions only 819 are classified as
verified transactions.
```{r}

```


```{r}
test %>% filter(isFraudPred != isFraud) %>% select(amount, oldbalanceOrg, newbalanceOrig,
                                                   newbalanceDest, newbalanceOrig) %>%
  ggpairs()
```

```{r}
sum(test %>% filter(isFraudPred != isFraud) %>% select(amount)) / 1000000
```
Amount of money that is lost due to transactions not handled correctly in Milliions


```{r}
summary(bst)
```

```{r}
# xgb.dump(bst, with_stats = T)
```

Plot the first tree of the model
```{r}
xgb.plot.tree(model = bst, trees = 1, show_node_id = TRUE)
```

